Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-07-08 introduction.tex
recovery intro Philippe
Petite modif intro

The latter case is related to a study that motivated the present work:
automated music transcription. Most representations of music are
essentially linear. This is true for audio files,
but also for widely used symbolic representations like MIDI.
Such representations ignore the hierarchical structures that frame the
conception of music, at least in the western area. These structures, on the other hand,
are present, either explicitly  or implicitly,
in music notation~\cite{Gould11Notation}: music scores are partitioned in measures, measures
in beats, and beats can be further recursively divided.
It follows that music events do not occur at arbitrary timestamps,
but respect a discrete division of the  timeline incurred by
these recursive divisions.
The \emph{transcription problem} takes
as input a linear representation (audio or MIDI) and aims at re-constructing
these structures
by mapping input events to this hierarchical rhythmic space.
It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990},
over an infinite alphabet of timed events.

@@

Other extensions of language models  %(automata and grammars)
help in dealing with non-determinism, by the computation of weight values.
With an ambiguous grammar, there may exist several derivations
(\emph{abstract syntax trees} -- AST) % \ie the result of an analyze.
yielding one input word. % structuring a word,
The association of one weight value %in semiring domain
to each AST permits to select a best one (or $n$ bests). % a fixed number of bests
% = ...a ranking of ASTs
This is roughly the principle of \emph{weighted parsing}
approaches~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}.
In \emph{weighted language models},
like \eg probabilistic context-free grammars % (CFG),
and weighted automata (\WA) \cite{Droste09handbook},
a weight value is associated to each transition rule, % production rule
and the rule's weights can be combined with an
associative product operator~$\otimes$ into the weight of an AST.
A second operator~$\oplus$, associative and commutative,
is moreover used to resolve the ambiguity raised by the existence
of several (in general exponentially many) AST
associated to a given input word.
Typically, $\oplus$ will select the best of two weight values.
%Intuitively,~$\oplus$ selects, or ranking, the syntax trees.
The weight domain, equipped with these two operators shall be, at minima,
a \emph{semiring} %$\Semiring$
where $\oplus$ can be extended to infinite sums,
%\cite{Eilenberg74automata}
such as the Viterbi semiring and the tropical min-plus algebra, see Figure~\ref{fig:semirings}.
\philippe{La figure 2 est citée avant la figure 1 mais apparait longtemps après. A corriger.}
%of domain $\mathbb{R}^+ \cup \{ +\infty\}$,
%where $\oplus$ is min and $\otimes$ is plus .
%by ranking
%making the weight domain a semiring.
%Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context
%% models represented as hypergraphs \cite{Huang05kbest}
%in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
%Generally based on dynamic programming, these algorithms rely on
%additional algebraic properties of~$\Semiring$.
%-- see \eg~\cite{Huang05kbest} for some NLP applications.
%The extraction of $n$ best list is useful
%the problem: quantitative parsing or symbolic parsing
%of parsing of words over infinite input alphabet.

>>>>

The latter case is related to a study that motivated the present work:
automated music transcription. Most representations of music are
essentially linear. This is true for audio files,
but also for widely used symbolic representations like MIDI.
Such representations ignore the hierarchical structures that frame the
conception of music, at least in the western area. These structures, on the other hand,
are present, either explicitly  or implicitly,
in music notation~\cite{Gould11Notation}: music scores are partitioned in measures, measures
in beats, and beats can be further recursively divided.
It follows that music events do not occur at arbitrary timestamps,
but respect a discrete partitioning of the  timeline incurred by
these recursive divisions.
The \emph{transcription problem} takes
as input a linear representation (audio or MIDI) and aims at re-constructing
these structures
by mapping input events to this hierarchical rhythmic space.
It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990},
over an infinite alphabet of timed events.

@@

Other extensions of language models  %(automata and grammars)
help in dealing with non-determinism, by the computation of weight values.
With an ambiguous grammar, there may exist several derivations
(\emph{abstract syntax trees} -- AST) % \ie the result of an analyze.
yielding one input word. % structuring a word,
The association of one weight value %in semiring domain
to each AST permits to select a best one (or $n$ bests). % a fixed number of bests
% = ...a ranking of ASTs
This is roughly the principle of \emph{weighted parsing}
approaches~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing}.
In \emph{weighted language models},
like \eg probabilistic context-free grammars % (CFG),
and weighted automata (\WA) \cite{Droste09handbook},
a weight value is associated to each transition rule, % production rule
and the rule's weights can be combined with an
associative product operator~$\otimes$ into the weight of an AST.
A second operator~$\oplus$, associative and commutative,
is moreover used to resolve the ambiguity raised by the existence
of several (in general exponentially many) AST
associated to a given input word.
Typically, $\oplus$ will select the best of two weight values.
%Intuitively,~$\oplus$ selects, or ranking, the syntax trees.
The weight domain, equipped with these two operators shall be, at minima,
a \emph{semiring} %$\Semiring$
where $\oplus$ can be extended to infinite sums,
%\cite{Eilenberg74automata}
such as the Viterbi semiring and the tropical min-plus algebra% % see Figure~\ref{fig:semirings}.
%of domain $\mathbb{R}^+ \cup \{ +\infty\}$,
%where $\oplus$ is min and $\otimes$ is plus .
%by ranking
%making the weight domain a semiring.
%Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context
%% models represented as hypergraphs \cite{Huang05kbest}
%in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
%Generally based on dynamic programming, these algorithms rely on
%additional algebraic properties of~$\Semiring$.
%-- see \eg~\cite{Huang05kbest} for some NLP applications.
%The extraction of $n$ best list is useful
%the problem: quantitative parsing or symbolic parsing
%of parsing of words over infinite input alphabet.
